---
title: "Capstone - Create Analysis Dataset"
subtitle: "Filter comments and replies posted after the Indonesian Presidential Election voting day and exclude videos with fewer than 10 total comments"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_knit$set(
  root.dir = "C:/Users/cheow/OneDrive - London School of Economics/Capstone/Data"
)
```


### Directory Management
```{r}
# Tell R my home folder
Sys.setenv(R_USER = "C:/Users/cheow/OneDrive - London School of Economics") 

# Specify directory 
wdir <- file.path(Sys.getenv("R_USER"), "Capstone")

# Sub-directory - to store data
ddir <- file.path(wdir, "Data")
if(!dir.exists(ddir)){
  dir.create(ddir)
}
```


### Install the packages
```{r}
# Check and install missing packages automatically
required_packages <- c("dplyr", "lubridate", "kableExtra")

for (pkg in required_packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    message("Installing missing package: ", pkg)
    install.packages(pkg)
  }
}
```


### Library
```{r}
library(dplyr)
library(lubridate)
library(kableExtra)
```


### Load full comments CSV file and reformat date for subsequent filtering steps 
```{r}
master_df_videos <- read.csv(
  file.path(ddir, "comments_all_playlists.csv"),
  stringsAsFactors = FALSE
) %>%
  mutate(
    published_at = as_datetime(published_at),
    date         = as.Date(published_at)
  )
```


### Filter comments (keep only comments/replies posted before 15 Feb 2024)
```{r}
# Voting date of Indonesian Presidential Election - 14 Feb 2024
analysis_df_videos <- master_df_videos %>%
  filter(date < as.Date("2024-02-15"))
```


### Verify 
```{r}
message("Total rows (master)    : ", nrow(master_df_videos))
message("Rows kept (pre-voting day)   : ", nrow(analysis_df_videos))
message("Rows dropped (post-voting day)    : ", nrow(master_df_videos) - nrow(analysis_df_videos))
message("Date range kept        : ", min(analysis_df_videos$date), " → ", max(analysis_df_videos$date))
```

### Filter out videos with less than 10 comments
```{r}
analysis_df_videos <- analysis_df_videos %>%
  group_by(video_id) %>%
  filter(n() >= 10) %>%
  ungroup()
```


### Save the dataset for all future analysis
```{r}
write.csv(analysis_df_videos,
          file.path(ddir, "videos_comments_analysis.csv"),
          row.names = FALSE)

message("Saved: ", nrow(analysis_df_videos), " rows → videos_comments_analysis.csv")
```


### Summary Statistics
```{r}

debate_labels <- c(
  debate1 = "First Presidential Debate",  debate2 = "Second Presidential Debate",
  debate3 = "Third Presidential Debate",  debate4 = "Fourth Presidential Debate",
  debate5 = "Fifth Presidential Debate"
)

videos_per_playlist <- analysis_df_videos %>%
  group_by(playlist_name) %>%
  summarise(n = n_distinct(video_id), .groups = "drop") %>%
  mutate(playlist_name = debate_labels[playlist_name])

data.frame(
  Statistic = c(
    "Total observations", "Top-level comments", "Replies",
    "Unique videos", "Unique users", "Date range (start)", "Date range (end)",
    paste0("", videos_per_playlist$playlist_name)
  ),
  Value = c(
    format(nrow(analysis_df_videos),                            big.mark = ""),
    format(sum(analysis_df_videos$comment_type == "top_level"), big.mark = ","),
    format(sum(analysis_df_videos$comment_type == "reply"),     big.mark = ","),
    format(n_distinct(analysis_df_videos$video_id),             big.mark = ","),
    format(n_distinct(analysis_df_videos$author_anon),          big.mark = ","),
    format(min(analysis_df_videos$date)), format(max(analysis_df_videos$date)),
    format(videos_per_playlist$n, big.mark = ",")
  )
) %>%
  kable(col.names = c("", ""), align = c("l", "r"),
        booktabs = TRUE, escape = FALSE,
        caption = "Descriptive Statistics of the collected YouTube comments") %>%
  kable_styling(latex_options = "hold_position", full_width = FALSE) %>%
  pack_rows("Sample Composition",  1, 3) %>%
  pack_rows("Coverage",            4, 7) %>%
  pack_rows("Videos per Playlist", 8, 7 + nrow(videos_per_playlist))

```



### Updated video-level statistics (videos used for analysis) 
```{r}
analysis_video_stat <- read.csv(file.path(ddir, "video_stats.csv"), stringsAsFactors = FALSE) %>%
  filter(video_id %in% unique(analysis_df_videos$video_id))

write.csv(analysis_video_stat, file.path(ddir, "analysis_video_stats.csv"), row.names = FALSE)
message("Stats saved: ", nrow(analysis_video_stat), " videos")
print(analysis_video_stat[, c("playlist_name", "video_title", "view_count", "comment_count", "like_count")])
```
