---
title: "Capstone - Comments Collection"
subtitle: "Code used to collect top-level comments and replies from Indonesian Presidential Debate livestream videos"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_knit$set(
  root.dir = "C:/Users/cheow/OneDrive - London School of Economics/Capstone/Data"
)
```


### Directory Management
```{r}
# Tell R my home folder
Sys.setenv(R_USER = "C:/Users/cheow/OneDrive - London School of Economics") 

# Specify directory 
wdir <- file.path(Sys.getenv("R_USER"), "Capstone")

# Sub-directory - to store data
ddir <- file.path(wdir, "Data")
if(!dir.exists(ddir)){
  dir.create(ddir)
}
```


### Install the packages
```{r}
# Check and install missing packages automatically
required_packages <- c("httr", "jsonlite", "dplyr", "lubridate", "keyring", "tuber")

for (pkg in required_packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    message("Installing missing package: ", pkg)
    install.packages(pkg)
  }
}
```


### Library
```{r}
library(httr)
library(jsonlite)
library(dplyr)
library(lubridate)
library(keyring)
library(tuber)
```


### Google Cloud - Project ID
```{r}
capstone_projectID <- "toxicity-analysis-486011"
```


### Retrieve API Key
```{r}
key_set("youtube_api_key")
api_key  <- key_get("youtube_api_key")
```


### Step 1 - Define presidential debate playlists from YouTube Indonesia
```{r}
# These playlists contain videos of the first to fifth Indonesian Presidential Debates compiled by YouTube Indonesia.
playlists <- c(
  debate1 = "PLF7VqZzmdYB5MZvpMnwe54ERI-VEJ6UfA",
  debate2 = "PLF7VqZzmdYB4OkGqJHH_YoNRo8VvOQozf",
  debate3 = "PLF7VqZzmdYB4DQ4FnJuw5IKIob-xE0RT8",
  debate4 = "PLF7VqZzmdYB7GSc-Kdnf0TH6SIu3nyc1l",
  debate5 = "PLF7VqZzmdYB7HxD-jnXIDmUO-yxVLnB6x"
)
```


### Step 2 - Create or load author map (for anonymisation of commenters)
```{r}
author_map_path_videos <- file.path(ddir, "author_map_videos.rds")

if (file.exists(author_map_path_videos)) {
  author_map_videos <- readRDS(author_map_path_videos)
  message("Loaded existing author_map_videos: ", length(author_map_videos), " entries.")
} else {
  author_map_videos <- list()
  message("No existing author_map_videos found. Starting fresh.")
}

```


### Step 3a - Define function: Fetch all video IDs from the defined playlists from Step 1
```{r}
fetch_playlist_videos <- function(playlist_id, api_key) {
  endpoint_pl <- "https://www.googleapis.com/youtube/v3/playlistItems"
  result      <- data.frame(video_id = character(), video_title = character(),
                             stringsAsFactors = FALSE)
  page_token  <- NULL

  repeat {
    params <- list(
      part       = "snippet",
      playlistId = playlist_id,
      maxResults = 50,
      key        = api_key
    )
    if (!is.null(page_token)) params$pageToken <- page_token

    resp <- GET(endpoint_pl, query = params)
    Sys.sleep(1)

    if (status_code(resp) != 200) {
      warning("Playlist fetch failed: ", playlist_id, " | Status: ", status_code(resp))
      break
    }

    data  <- fromJSON(content(resp, as = "text", encoding = "UTF-8"), flatten = TRUE)
    items <- data$items

    if (!is.null(items) && nrow(items) > 0) {
      batch <- data.frame(
        video_id    = items$snippet.resourceId.videoId,
        video_title = items$snippet.title,
        stringsAsFactors = FALSE
      )
      result <- bind_rows(result, batch)
    }

    page_token <- data$nextPageToken
    if (is.null(page_token)) break
  }

  message("Found ", nrow(result), " videos in playlist: ", playlist_id)
  return(result)
}

```


### Step 3b - Define function: Fetch video stats
```{r}
fetch_video_stats <- function(video_ids, api_key) {
  endpoint_videos <- "https://www.googleapis.com/youtube/v3/videos"
  result          <- data.frame()

  # YouTube API allows max 50 IDs per call — batch accordingly
  batches <- split(video_ids, ceiling(seq_along(video_ids) / 50))

  for (batch in batches) {
    params <- list(
      part = "snippet,statistics",
      id   = paste(batch, collapse = ","),
      key  = api_key
    )

    resp <- GET(endpoint_videos, query = params)
    Sys.sleep(1)

    if (status_code(resp) != 200) {
      warning("Video stats fetch failed | Status: ", status_code(resp))
      next
    }

    data  <- fromJSON(content(resp, as = "text", encoding = "UTF-8"), flatten = TRUE)
    items <- data$items

    if (!is.null(items) && nrow(items) > 0) {
      batch_df <- data.frame(
        video_id      = items$id,
        video_title   = items$snippet.title,
        view_count    = as.numeric(items$statistics.viewCount),
        comment_count = as.numeric(items$statistics.commentCount),
        like_count    = as.numeric(items$statistics.likeCount),
        stringsAsFactors = FALSE
      )
      result <- bind_rows(result, batch_df)
    }
  }

  return(result)
}

```


### Step 3c - Define helper functions: Collect comments for a single video
```{r}

# Helper function to prevent comments collection error from arising due to deleted accounts
safe_author <- function(a) {
  if (is.null(a) || length(a) == 0 || (length(a) == 1 && is.na(a))) return(NA_character_)
  val <- tryCatch(as.character(a[[1]]), error = function(e) NA_character_)
  if (is.na(val) || nchar(trimws(val)) == 0) return(NA_character_)
  val
}

# Helper function to prevent data.frame() row mismatch error due to NULL, character(0), NA, 
# or length > 1 values from the API, coercing any input to a guaranteed length-1 scalar.
safe_val <- function(x, default = NA_character_) {
  if (is.null(x) || length(x) == 0 || (length(x) == 1 && is.na(x))) return(default)
  val <- tryCatch(x[[1]], error = function(e) default)
  if (is.null(val) || length(val) == 0 || is.na(val)) return(default)
  if (length(val) > 1) val <- val[[1]]
  val
}

# Helper function to retry comments collection due to error in row build error, or network limit failure
get_with_retry <- function(url, params, max_retries = 3) {
  for (attempt in seq_len(max_retries)) {
    resp <- GET(url, query = params)
    Sys.sleep(1)

    if (status_code(resp) == 200) {
      return(content(resp, as = "parsed"))   # key change: list not flattened data frame
    }
    if (status_code(resp) == 403) {
      message("  Access forbidden (403) — skipping permanently.")
      return(NULL)
    }
    if (status_code(resp) == 429) {
      wait <- attempt * 10
      message("  Rate limited (429) — waiting ", wait, "s before retry ", attempt, "/", max_retries)
      Sys.sleep(wait)
    } else {
      message("  Attempt ", attempt, " failed (", status_code(resp), ") — retrying...")
      Sys.sleep(attempt * 2)
    }
  }
  message("  All ", max_retries, " attempts failed — skipping.")
  NULL
}

```


### Step 3d - Define function: Collect comments for a single video
```{r}
collect_comments_for_video <- function(video_id, playlist_id, playlist_name,
                                        video_title, api_key) {
  endpoint_threads <- "https://www.googleapis.com/youtube/v3/commentThreads"
  endpoint_replies <- "https://www.googleapis.com/youtube/v3/comments"

  all_rows       <- list()
  page_token     <- NULL
  page_num       <- 0
  total_toplevel <- 0
  total_replies  <- 0
  ts             <- function() format(Sys.time(), "[%H:%M:%S]")

  message(ts(), " ┌─────────────────────────────────────────")
  message(ts(), " │ VIDEO : ", video_title)
  message(ts(), " │ ID    : ", video_id)
  message(ts(), " │ LIST  : ", playlist_name)
  message(ts(), " └─────────────────────────────────────────")

  # ── Top-level comment threads ──────────────────────────────────────────────
  repeat {
    params <- list(
      part       = "snippet",
      videoId    = video_id,
      maxResults = 100,
      textFormat = "plainText",
      key        = api_key
    )
    if (!is.null(page_token)) params$pageToken <- page_token

    page_num <- page_num + 1
    message(ts(), "   PAGE ", page_num, " — fetching...")

    data <- get_with_retry(endpoint_threads, params)
    if (is.null(data)) {
      message(ts(), "   ✖ Skipping video (forbidden or all retries failed): ", video_id)
      return(NULL)
    }

    items <- data$items
    message(ts(), "   PAGE ", page_num, " — ", length(items), " comments found")

    for (item in items) {

      snip   <- item$snippet$topLevelComment$snippet   # list-style access
      cid    <- item$snippet$topLevelComment$id
      nr     <- item$snippet$totalReplyCount
      author <- safe_author(snip$authorDisplayName)

      # Anonymise author
      if (is.na(author) || trimws(author) == "") {
        author_anon <- "Unknown"
      } else if (author %in% names(author_map_videos)) {
        author_anon <- author_map_videos[[author]]
      } else {
        author_anon          <- paste0("User_", length(author_map_videos) + 1)
        author_map_videos[[author]] <<- author_anon
      }

      # Retry-then-skip: top-level comment row
      row_built <- FALSE
      for (attempt in 1:3) {
        tryCatch({
          all_rows[[length(all_rows) + 1]] <- data.frame(
            playlist_id   = playlist_id,
            playlist_name = playlist_name,
            video_id      = video_id,
            video_title   = video_title,
            comment_id    = safe_val(cid),
            parent_id     = NA_character_,
            comment_type  = "top_level",
            author_anon   = author_anon,
            comment       = safe_val(snip$textDisplay),
            published_at  = safe_val(snip$publishedAt),
            like_count    = as.numeric(safe_val(snip$likeCount, 0)),
            reply_count   = as.numeric(safe_val(nr, 0)),
            stringsAsFactors = FALSE
          )
          row_built      <- TRUE
          total_toplevel <- total_toplevel + 1
        }, error = function(e) {
          message(ts(), "     ✖ Attempt ", attempt, " failed [comment]: ", e$message)
        })
        if (row_built) break
      }
      if (!row_built) message(ts(), "     ✖ Permanently skipped a top-level comment in ", video_id)

      # ── Replies ──────────────────────────────────────────────────────────────
      if (!is.null(nr) && !is.na(nr) && nr > 0) {
        reply_page_token <- NULL
        reply_page_num   <- 0

        repeat {
          reply_params <- list(
            part       = "snippet",
            parentId   = cid,
            maxResults = 100,
            textFormat = "plainText",
            key        = api_key
          )
          if (!is.null(reply_page_token)) reply_params$pageToken <- reply_page_token

          reply_page_num <- reply_page_num + 1
          reply_data     <- get_with_retry(endpoint_replies, reply_params)
          if (is.null(reply_data)) break

          reply_items <- reply_data$items
          message(ts(), "     ↳ replies page ", reply_page_num,
                  " — ", length(reply_items), " retrieved")

          for (reply in reply_items) {

            r_snip   <- reply$snippet                  # ← list-style access
            r_author <- safe_author(r_snip$authorDisplayName)

            # Anonymise reply author
            if (is.na(r_author) || trimws(r_author) == "") {
              r_anon <- "Unknown"
            } else if (r_author %in% names(author_map_videos)) {
              r_anon <- author_map_videos[[r_author]]
            } else {
              r_anon                 <- paste0("User_", length(author_map_videos) + 1)
              author_map_videos[[r_author]] <<- r_anon
            }

            # Retry-then-skip: reply row
            row_built <- FALSE
            for (attempt in 1:3) {
              tryCatch({
                all_rows[[length(all_rows) + 1]] <- data.frame(
                  playlist_id   = playlist_id,
                  playlist_name = playlist_name,
                  video_id      = video_id,
                  video_title   = video_title,
                  comment_id    = safe_val(reply$id),
                  parent_id     = safe_val(cid),
                  comment_type  = "reply",
                  author_anon   = r_anon,
                  comment       = safe_val(r_snip$textDisplay),
                  published_at  = safe_val(r_snip$publishedAt),
                  like_count    = as.numeric(safe_val(r_snip$likeCount, 0)),
                  reply_count   = 0L,
                  stringsAsFactors = FALSE
                )
                row_built     <- TRUE
                total_replies <- total_replies + 1
              }, error = function(e) {
                message(ts(), "       ✖ Attempt ", attempt, " failed [reply]: ", e$message)
              })
              if (row_built) break
            }
            if (!row_built) message(ts(), "       ✖ Permanently skipped a reply in ", video_id)
          }

          reply_page_token <- reply_data$nextPageToken
          if (is.null(reply_page_token)) break
        }
      }
    }

    page_token <- data$nextPageToken
    if (is.null(page_token)) break
  }

  if (length(all_rows) == 0) {
    message(ts(), "   ✖ No comments collected for: ", video_id)
    return(NULL)
  }

  message(ts(), " ✔ DONE  │ Pages: ", page_num,
          " │ Top-level: ", total_toplevel,
          " │ Replies: ", total_replies,
          " │ Total: ", total_toplevel + total_replies)
  message()

  bind_rows(all_rows) %>%
    mutate(
      published_at = as_datetime(published_at),
      date         = as.Date(published_at),
      time         = format(published_at, "%H:%M:%S")
    )
}
 
```


### Step 4 - Master Collection Loop
```{r}
all_videos_data <- list()
all_stats       <- list()

for (pl_name in names(playlists)) {
  pl_id <- playlists[[pl_name]]
  message("\n=== Playlist: ", pl_name, " (", pl_id, ") ===")

  pl_videos <- fetch_playlist_videos(pl_id, api_key)
  if (nrow(pl_videos) == 0) { message("  No videos found. Skipping."); next }

  # Fetch and store video stats for this playlist
  stats_df               <- fetch_video_stats(pl_videos$video_id, api_key)
  stats_df$playlist_name <- pl_name
  stats_df$playlist_id   <- pl_id
  all_stats[[pl_name]]   <- stats_df

  # Collect comments for each video
  for (v in seq_len(nrow(pl_videos))) {
    vid_id    <- pl_videos$video_id[v]
    vid_title <- pl_videos$video_title[v]

    # Skip if already collected (crash recovery)
    out_path <- file.path(ddir, paste0("comments_", pl_name, "_", vid_id, ".csv"))
    if (file.exists(out_path)) {
      message("  Already collected, skipping: ", vid_id)
      all_videos_data[[paste0(pl_name, "_", vid_id)]] <- read.csv(out_path, stringsAsFactors = FALSE)
      next
    }

    message("  Collecting: ", vid_title, " (", vid_id, ")")
    vid_df <- collect_comments_for_video(vid_id, pl_id, pl_name, vid_title, api_key)

    if (!is.null(vid_df) && nrow(vid_df) > 0) {
      write.csv(vid_df, out_path, row.names = FALSE)
      saveRDS(author_map_videos, author_map_path_videos)  # save after every video
      message("    Saved: ", nrow(vid_df), " rows → ", basename(out_path))
      all_videos_data[[paste0(pl_name, "_", vid_id)]] <- vid_df
    } else {
      message("    No comments returned for: ", vid_id)
    }
  }
}

```


### Combine and save all video comments + Video Stats
```{r}
# Standardise published_at across all collected data frames before combining
all_videos_data <- lapply(all_videos_data, function(df) {
  df$published_at <- as_datetime(df$published_at)
  df$date         <- as.Date(df$published_at)
  df$time         <- format(df$published_at, "%H:%M:%S")
  df
})

# Save master comments CSV
if (length(all_videos_data) > 0) {
  master_df <- bind_rows(all_videos_data)
  write.csv(master_df, file.path(ddir, "comments_all_playlists.csv"), row.names = FALSE)
  message("\nDone. Master CSV: ", nrow(master_df), " rows, ",
          n_distinct(master_df$video_id), " videos, ",
          n_distinct(master_df$playlist_id), " playlists.")
}

# Save video stats CSV
if (length(all_stats) > 0) {
  stats_master <- bind_rows(all_stats)
  write.csv(stats_master, file.path(ddir, "video_stats.csv"), row.names = FALSE)
  message("Stats saved: ", nrow(stats_master), " videos")
  print(stats_master[, c("playlist_name", "video_title", "view_count", "comment_count", "like_count")])
}

```

